---
title: "Assignment 8"
format: html
editor: visual
author: Agata Leszczak
---

# Visualization Improvement Report

## 1. British Museum Artifacts by Country of Origin

### Original Problem

The original visualization had scaling issues where columns weren't proportional to their values (e.g., a value of 29,280 appeared about 1/3 the size of 625,371).

### Improved Visualization

```{r}
library(ggplot2)
library(dplyr) 

# Simulate data for British Museum artifacts by country 
set.seed(123)
countries <- c("United Kingdom", "Egypt", "Greece", "Italy", "France",                 "China", "Iraq", "Turkey", "India", "Germany", "Japan") 
stolen_things_counts <- c(625371, 29280, 18479, 16885, 15983, 14937,                       12483, 11289, 10045, 8942, 75000)  # Approximate values  
data <- data.frame(   Country = factor(countries, levels = countries[order(stolen_things_counts, decreasing = TRUE)]),   Count = stolen_things_counts )  # Create improved bar plot 
p1 <- ggplot(data, aes(x = Country, y = Count)) +   
  geom_bar(stat = "identity", fill = "#1f77b4", width = 0.7) +  
  geom_text(aes(label = format(Count, big.mark = ",")),              
            vjust = -0.3, size = 3) +  
  labs(     title = "British Museum Artifacts by Country of Origin",    
            subtitle = "Properly scaled visualization with clear labels",   
            x = "Country of Origin",   
            y = "Number of Artifacts"   ) + 
  theme_minimal() + 
  theme(     axis.text.x = element_text(angle = 45, 
                                        hjust = 1, size = 9),   
             axis.text.y = element_text(size = 9),   
             plot.title = element_text(size = 16, face = "bold"), 
             plot.subtitle = element_text(size = 12, color = "gray40"),  
             panel.grid.major.x = element_blank(),    
             panel.grid.minor.y = element_blank()   ) + 
  scale_y_continuous(     labels = scales::comma,  
                          expand = expansion(mult = c(0, 0.1))   ) 
print(p1)
```

**Improvements Made:**

-   Proper scaling of bars to accurately represent values

-   Clear labeling of exact values above each bar

-   Rotated x-axis labels for better readability

-   Clean, minimal theme with appropriate grid lines

-   Thousands separators in y-axis labels

## 2. American Travel to Europe by Month and Country

### Original Problem

The original graph had an inappropriate scale that made it impossible to determine actual travel numbers, with all values compressed visually.

### Improved Visualization

I added all the months just because and I also assumed that the missing/non visible Us to Europe was just normal, random numbers similar to the others.

```{r}
# Simulate data for American travel to Europe
set.seed(456)
months <- month.name 
countries_travel <- c("US to Europe", "Canada to Europe", "Canada from US")  # Create realistic monthly travel data (in thousands)
travel_data <- expand.grid(   Month = factor(months, levels = months),   Country = countries_travel )  # Add simulated travel counts  
travel_data$Travelers <- c(  
  c(450, 420, 480, 20, 580, 50, 720, 100, 620, 550, 480, 430),    
  c(380, 350, 420, 500, 580, 640, 690, 670, 590, 510, 440, 400),   
  
  c(320, 310, 380, 220, 500, 620, 150, 730, 410, 480, 370, 330) ) * 1000 
plot2 <- ggplot(travel_data, aes(x = Month, y = Travelers, fill = Country)) +  
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +  
  geom_text(     aes(label = format(round(Travelers/1000, 0), big.mark = ",")),   
                 position = position_dodge(width = 0.8),    
                 vjust = -0.3, size = 2.8, angle = 0   ) +   
  labs(     title = "Q1 2025 Travel Flows",    
            subtitle = "Monthly travel patterns with appropriate scale (values in thousands)",     x = "Month",     
            y = "Number of American Travelers",  
            fill = "Destination Country : "   ) +   
  theme_minimal() +  
  theme(     axis.text.x = element_text(angle = 45, hjust = 1, size = 9),    
             axis.text.y = element_text(size = 9),    
             plot.title = element_text(size = 16, face = "bold"),    
             plot.subtitle = element_text(size = 10, color = "gray40"),    
             legend.position = "top",    
             panel.grid.major.x = element_blank()   ) +   
  scale_fill_manual(values = c("#1f77b4", "#ff7f0e", "#2ca02c")) +   
  scale_y_continuous(     labels = function(x) format(x/1000, big.mark = ","),     expand = expansion(mult = c(0, 0.15))   )  

print(plot2)
```

**Improvements Made:**

-   Appropriate y-axis scale that shows meaningful variation

-   Clear grouping by month with country differentiation

-   Direct labeling of approximate values

-   Color coding for easy country identification

-   Proper title and axis labels

## 
