---
title: "Assignment 7 Agata Leszczak"
format: html
editor: visual
---

## Task1

**Task 1:** Load the “iris” dataset and transform it to longer.

-   tip: you should get 3 columns Species \| Trait \| Size

1.  I need to load all the libraries and then view the files

```{r}
library(tidyr)
library(dplyr)
```

```{r}
data(iris)
head(iris)
```

```{r}
iris_long <- iris %>%
  pivot_longer(
    cols = -Species, 
    names_to = "Trait", 
    values_to = "Size"  
  )

```

I took all measurement columns ( except Species) and stacked them into a single column "Trait", with the numerical values placed in "Size".

```{r}
head(iris_long)
```

## Task 2:

Answer the following questions.

```         
tip: you can answer each question “piping” only 2-3 commands
```

Task 2.1: What is the mean length and width of the sepal and petal of each species?

For that I grouped the data by species and traits and then did the mean of the size column. I addded the .group arg since without it it wouldn't work at all. I switched to wider format for bettervisibility

```         
```

```{r}
mean_sizes <- iris %>%
  group_by(Species) %>%
  summarise(
    Sepal.Length = mean(Sepal.Length),
    Sepal.Width  = mean(Sepal.Width),
    Petal.Length = mean(Petal.Length),
    Petal.Width  = mean(Petal.Width)
  )


```

```{r}
head(mean_sizes)
```

Task 2.2: What is the largest and widest, petal and sepal, of virginica and versicolor species?

```{r}
largest_values <- iris_long %>%
  filter(Species %in% c("virginica", "versicolor")) %>%
  group_by(Species, Trait) %>%
  summarise(
    Max_Size = max(Size),
    Min_Size = min(Size),
    .groups = "drop"
  ) %>%
  arrange(Species, Trait)

largest_values
```

Task 2.3: For each species, for sepal and petal width and length, how many observations are above the mean?\*

```         
tip: adding the means as column might help
```

```{r}

above_mean <- iris_long %>%
  group_by(Species, Trait) %>% 
  summarise(
    Average_Size = mean(Size),
    Count_Above_Mean = sum(Size > Average_Size), .groups = "drop"
  )

above_mean
```

## Task 3:

Deliver a link to GitHub containing the a Quarto file and the compiled HTML. It should contain explanations of what each chunk does. If you have private repository, remember to send me invitation so I can access it .

```         
```
